---
title: "P8105 Homework 6"
output: github_document
date: "December 3rd, 2022"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  fig.dim = c(12, 7))

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Part 1: Central Park Weather Data

For this problem, we’ll use the 2017 Central Park weather data from the `rnoaa` package. The code chunk below will download these data.

```{r load noaa}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
crossval_df = weather_df %>% 
  modelr::bootstrap(n = 100) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  group_by(term) %>% 
  summarize(boot_se = sd(estimate))
```

## Part 2: Washington Post Homicide Data

### Load and clean data

The Washington Post has gathered data on homicides in 50 large U.S. cities. Let's load the raw homicide data and inspect it.

```{r load homicide data}
homicide_data = read_csv("data/homicide-data.csv") %>% 
  janitor::clean_names()
homicide_data
```

There are `r nrow(homicide_data)` observations of `r ncol(homicide_data)` variables containing information about homicides in 50 cities across the US. The data contains the date, victim information (name, age, sex, race), location (city, state, latitude, longitude), and status of the case.

Next, let's tidy the data by creating a `city_state` variable, and omitting Dallas, Phoenix, Kansas City, and Tulsa, AL (a data entry error). We will also convert `victim_age` to a numeric variable, and keep observations only where the victim's race is White or Black.

```{r clean data}
homicide_data = homicide_data %>% 
  mutate(
    state = str_to_upper(state),
    city_state = str_c(city, state, sep = ", "), 
    victim_age = as.numeric(victim_age),
    resolved = as.numeric(disposition == "Closed by arrest")) %>% 
  filter(city_state != "Dallas, TX" & 
           city_state != "Phoenix, AZ" & 
           city_state != "Kansas City, MO" & 
           city_state != "Tulsa, AL", 
         victim_race == "White" | victim_race == "Black")
```

There are now `r nrow(homicide_data)` observations of `r ncol(homicide_data)` variables in our cleaned dataset containing information about homicides in 47 cities across the US.

### Modelling homicide data

Now, using purrr::map, list columns, and unnest, we will run `glm` for each of the cities in the dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 

The resulting code creates a dataframe with estimated adjusted ORs and 95% CIs for each city, comparing the odds of solving a homicide among male victims compared to females.

```{r map glm}
OR_df = homicide_data %>% 
  nest(data = -city_state) %>% 
  mutate(models = map(data, ~ glm(resolved ~ victim_age + victim_sex + victim_race, 
                                 data = ., family = "binomial")), 
         results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  mutate(OR = exp(estimate),
         CIL = exp(estimate - 1.96*std.error),
         CIU = exp(estimate + 1.96*std.error)) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, term, log_OR = estimate, OR, CIL, CIU, p.value)
```

Next, we will plot the estimated ORs and CIs for each city.

```{r plot ORs}
OR_df %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR, colour = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CIL, ymax = CIU, width = .3)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        legend.position = "none") + 
  labs(
    x = "Location",
    y = "Adjusted odds ratio",
    title = "Effect of male vs. female victim sex on the odds of solving homicides across major U.S. cities, 2007-2017",
    caption = paste0(
          "Error bars represent the 95% confidence interval.",
          "\nSource: The Washington Post."))
```

We can see that across a majority of cities, the ORs are less than 1, indicating a decreased odds of resolving a homicide when the victim is male vs. female. This effect is the greatest for New York, Baton Rouge, and Omaha. Many cities also have a non-significant association between victim sex and resolving homicides, indicated by the 95% CI for the adjusted OR crossing the null value of 1. 

On the other hand, there is an increased odds of solving a homicide when the victim is male vs. female in Albequerque, Stockton, and Fresno. However, for the 3 cities with the highest adjusted ORs, the 95% CI crosses the null value of 1, indicating there may be a high variability in these estimates rather than an effect of victim sex on solving homicides.



